---
title: "Introduction to fscache package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*fscache* helps handle a user file system cache for an application or package.

Let us suppose we have an application called `myapp` for which we want to save
text content on disk in order to reload it later.
With *fscache* we will create a `Cache` object that will place our content in a
standard location, according to the poperating system we are running on, and
help us manage it (save, load, delete, place in sub-folders, etc).

## Initializing the cache

First, start by loading the package:
```{r setup}
library(fscache)
```

Then we create a `Cache` instance:
```{r}
my.folder.cache <- file.path(tempdir(), 'my.cache') # We would usually use here a relative path like 'myapp'
cache <- Cache$new(my.folder.cache)
```
For this vignette, we use a temporary folder as our cache folder.
Since the temporary folder is an absolute path, the user HOME folder will not
be used.
However, usually, in practice, we give the name of our application as the cache
folder to create.
The folder will be created in the user space, inside the standard path defined
for the operating system, given by the `tools`  package.

If we look at the `cache` object, we can see the exact location of the created
cache folder:
```{r}
cache
```
We also see that the cache is both *readable* and *writable* by default.
However we may block either read access or write access for our cache, using
the `setReadable()` and `setWritable()` methods.
It is sometimes useuful for testing purposes.

The full path to the cache folder is also accessible through the `getFolder()`
method:
```{r}
cache$getFolder()
```

## Saving and loading content

Let us save the following text contents into the cache:
```{r}
x <- c('abc', 'def', 'ghi')
```

For that we use the `saveContents()` method:
```{r}
cache$saveContents(x, c('a.txt', 'b.txt', 'c.txt'))
```
We have saved our three strings into three files inside the cache folder.
Since all three files have the same extension, we may have used the `suffix`
parameters to avoid repeating the extension in the filenames:
```{r}
cache$saveContents(x, c('a', 'b', 'c'), suffix='.txt')
```
Note that *fscache* does not complain that the files already exist. It
overrides them silently.

We can list the existing files inside the cache folder with the following
command:
```{r}
cache$globPaths(suffix='.txt')
```

Loading the contents from cached files is done with the `loadContents()`
method, which returns a `character` vector with names set with the file paths:
```{r}
cache$loadContents(c('a', 'b', 'c'), suffix='.txt')
```
Note that if we try to load the content from non-existing file, a `NA` value
will be returned:
```{r}
cache$loadContents(c('a', 'b', 'c', 'd'), suffix='.txt')
```

Prior to loading contents, we may test the existence of files inside the cache
folder, with the `pathsExist()` method:
```{r}
cache$pathsExist(c('a', 'b', 'c', 'd'), suffix='.txt')
```

## Using sub-folders

Instead of putting them directly inside the cache folder, files may be put
inside sub-folders to organize them.

Here we put our contents into three new files inside a sub-folder named
`"mysub"` that will be automatically created:
```{r}
cache$saveContents(x, c('x', 'y', 'z'), suffix='.txt', sub.folder='mysub')
```
Let us look a the content of our sub-folder:
```{r}
cache$globPaths(sub.folder='mysub')
```

## Deleting files

Files may be deleted from the cache, as in the following command:
```{r}
cache$delPaths(c('a', 'c'), suffix='.txt')
```
The remaining files are:
```{r}
cache$globPaths()
```

For deleting files inside a sub-folder, use the `sub.folder` parameter:
```{r}
cache$delPaths('y', suffix='.txt', sub.folder='mysub')
```

For deleting a sub-folder with all its content, use the `delFolder()` method:
```{r}
cache$delFolder('mysub')
```

## Copying or moving files into the cache

It is also possible to move or copy files that reside outside the cache folder,
into the cache folder.

Let us create a file:
```{r}
my.file <- tempfile('my.file', fileext='.txt')
cat('My text content.', file=my.file)
my.file
```

To move it into the cache, we use the `importFiles()` method:
```{r}
cache$importFiles(my.file, action='move')
```
For copying, we would have set `action` to `"copy"`.

Here the new list of cached files:
```{r}
cache$globPaths()
```

## Erasing the cache

If we wish to delete the whole cache folder with all its content, we have to
run the following command:
```{r}
cache$erase()
```
